---
title: "Analysis covid-19 preprocessing"
output: html_notebook
---

## Data specificity
- paired samples
- time course experiments
- 4 conditions(6 comparisons)

## Data input and output
Inputs : 
Proteomics raw data is submitted with uniprot accession as row and samples as columns.
Annotation file with as columns 1) Uniprot Accession 2) Gene name 

Ouputs :  
1)  PCA 1/ 2, UMAP, PLS-DA, Distribution after norm
2)  Heatmap, pathways, figure with pathways
3)  Scatter plot, Correlation matrix, pathways, figure pathways


## Outline 
### Pre-processing
Count missing values / PCA 1
Normalization ( NormalizerDE, selection based on report)
Plotting Distribution / Dimension reduction : PCA 2 / UMAP / PLS-DA

### Multiple comparisons
LIMMA : 6 comparisons (UNINF/T1 / T2 / T3) paired samples
Heatmap (hierarchical clustering) with proteins significant in at least one infected sample compared to control (pvalue  < 0.05).
Cytoscape Network → Reactome on obtained clusters

### Correlation viral/ host proteins
Scatter plot log2 (viral_protein_inf_sample/viral_prot_control) vs time after infection
Calculate averaged profile of viral proteins
Pearson correlation host proteins/ averaged profile of all viral proteins => select FDR < 0.01
Cytoscape Network → biological process GO terms (plot network)


### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Corona_omics/")) 
```

### load packages
```{r}
library(ggplot2)
library(limma)
library(dplyr)
library(xlsx)
library(RColorBrewer)
library(gplots)
library(ggrepel)
library(knitr)
library(ggfortify)
library(NormalyzerDE)
opts_chunk$set(echo = TRUE)
options(bitmapType="cairo")
```

### load own functions
```{r}
source("src/usefull_functions.R")
```

```{r}
name_ana <- "covid-19_proteomics"

nb_samples <- 3

# define conditions
t0 <- "UNINF"
t24 <- "H24PI"
t48 <- "H48PI"
t72 <- "H72PI"

# define associated colors
c_t0 <- "#B1A78C"
c_t24 <- "#687E8A"
c_t48 <- "#41607D"
c_t72 <- "#09253B"
  
condition <- as.factor(c(rep(t0, nb_samples), rep(t24, nb_samples), rep(t48, nb_samples), rep(t72, nb_samples)))
colors <- as.factor(c(rep(c_t0, nb_samples), rep(c_t24, nb_samples), rep(c_t48, nb_samples), rep(c_t72, nb_samples)))
gg_colors <- c(c_t24,c_t48, c_t72, c_t0)

input <- "data/Proteomics/20200411_L1_Huh7_CoV_TMTPro_Fractions.csv"
```

```{r}
dir.create(paste0("processing/", name_ana))
```

```{r}
dir.create(paste0("results/", name_ana))
dir.create(paste0("results/", name_ana, "/figures"))
dir.create(paste0("results/", name_ana, "/DGE"))
dir.create(paste0("results/", name_ana, "/Time_series"))
dir.create(paste0("results/", name_ana, "/GSEA"))
```

## 1) Pre-processing
```{r}
data <- read.csv(input)
annotationInfo <- select(data, Accession, Gene.Symbol)
names(annotationInfo)[2] <- "Gene_name"
```

## select Abundances columns
```{r}
data <- select(data, Accession, starts_with("Abundance.."))

rownames(data) <- data$Accession
data$Accession <- NULL

remov_row <- data[rowSums(is.na(data)) == ncol(data), ]
 
data <- data[rowSums(is.na(data)) != ncol(data), ] ## remove empty lines

## order samples
data <- data[,c(1, 3, 2, 4, 6, 5, 9, 8, 7, 10, 11, 12)]
```

## rename samples

```{r}
new_names <- c(paste0(t0, "_",1:3), paste0(t24,"_",1:3), paste0(t48,"_",1:3), paste0(t72,"_",1:3))

information <- cbind(colnames(data), new_names)

colnames(data) <- new_names

save_file_csv(annotationInfo, "annotationInfo", paste0("processing/", name_ana, "/"))
save_file_csv(data, "data_corona_filt", paste0("processing/", name_ana, "/"))
save_file_csv(information, "names_samples", paste0("processing/", name_ana, "/"))
```
### check for missing values
```{r}
counts <- 1:nrow(data) # create vector of appropriate length
for(i in 1:nrow(data)){
    # TRUE will be coerced to 1 for the summing
    counts[i] <- sum(is.na(data[i,]))
}
table(counts) # create the summary

save_file_csv(table(counts), "data_corona_missing_values", paste0("processing/", name_ana, "/"))
```
### distribution
```{r}
data_2 <-as.matrix(data)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
save_figure(data_2, "distribution_", 10, 10)
```

### PCA before norm
```{r}
pca_data <- data.frame(group = condition, 
                       t(as.matrix(data)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```

```{r}
a <- autoplot(pca, data = pca_data, colour = 'group', size= 6)+ scale_color_manual(values = gg_colors)
print(a)
save_figure(a, "pca_before_norm_", 10, 10)
```

```{r}
a <- autoplot(pca, data = pca_data, colour = 'group', size= 6, label = TRUE, shape = FALSE)+ scale_color_manual(values = gg_colors)
#print(a)
save_figure(a, "pca_before_norm_labelled", 10, 10)
```


### Quantile normalization
```{r}
path_data <- make_data_matrix(data, name_ana)
path_design <-make_design_matrix(data, name_ana, condition)
outDir <- "processing/normalization"
```

```{r}
#normalyzer(jobName=name_ana, designPath=path_design, dataPath=path_data, outputDir=outDir)
```

```{r}
path_results_norm <- paste0(outDir, "/", name_ana,"/Quantile-normalized.txt")
data_norm <- read.delim(path_results_norm)
rownames(data_norm) <- rownames(data)

path_results <- paste0("processing/norm_files/", name_ana, "_norm_file.txt")

write.table(data_norm, file = path_results, 
            sep = "\t", row.names = TRUE, col.names = NA)
```


### distribution after normalization
```{r}
data_2 <-as.matrix(data_norm)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
save_figure(data_2, "distribution_after_norm_", 10, 10)
```

### PCA after normalization
```{r}
pca_data <- data.frame(group = condition, t(as.matrix(data_norm)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```

```{r}
b <- autoplot(pca, data = pca_data, colour = 'group', size= 6)+ scale_color_manual(values = gg_colors)
print(b)
save_figure(b, "pca_after_norm_", 10, 10)
```
