---
title: "Time serie proteomics"
output: html_notebook
---


### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Corona_omics/")) 
```

### load packages
```{r}
library(psych)
library(GGally)
library(ggplot2)
library(limma)
library(dplyr)
library(xlsx)
library(RColorBrewer)
library(gplots)
library(ggrepel)
library(knitr)
library(ggfortify)
#library(NormalyzerDE)
opts_chunk$set(echo = TRUE)
options(bitmapType="cairo")
```

### load own functions
```{r}
source("src/usefull_functions.R")
```
```{r}
name_ana <- "covid-19_proteomics"

nb_samples <- 3

# define conditions
t0 <- "UNINF"
t24 <- "H24PI"
t48 <- "H48PI"
t72 <- "H72PI"

# define associated colors
c_t0 <- "#B1A78C"
c_t24 <- "#687E8A"
c_t48 <- "#41607D"
c_t72 <- "#09253B"
  
condition <- as.factor(c(rep(t0, nb_samples), rep(t24, nb_samples), rep(t48, nb_samples), rep(t72, nb_samples)))
time <- c(rep(0, nb_samples), rep(24, nb_samples), rep(48, nb_samples), rep(72, nb_samples))
colors <- as.factor(c(rep(c_t0, nb_samples), rep(c_t24, nb_samples), rep(c_t48, nb_samples), rep(c_t72, nb_samples)))
gg_colors <- c(c_t24,c_t48, c_t72, c_t0)

input <- read.delim("processing/norm_files/covid-19_proteomics_norm_file.txt")
rownames(input) <- input$X
input$X <- NULL

annotationInfo <- read.csv("processing/covid-19_proteomics/covid-19_proteomics_annotationInfo.csv")
annotationInfo$Gene_name <- gsub(";.*","",annotationInfo$Gene_name)
annotationInfo$X <- NULL
```
```{r}
source("src/usefull_functions.R")
```

```{r}
input_2 <- select(input, contains("24"),  contains("48"), contains("72"))
```

```{r}
input_2 <- select(input_2, contains("_1"),  contains("_2"), contains("_3"))
```

```{r}
names(input_2)
```


```{r}
assay <- c(rep("Sample..1", 3),
           rep("Sample..2", 3),
           rep("Sample..3", 3))

time.grp <- rep(c(24, 48, 72), 3)
```

## 1) LIMMA

## set levels
```{r}
groups <- as.factor(assay)
design <- model.matrix( ~0 + groups + time.grp)
```

## fit model
```{r}
fit <- lmFit(input_2, design)
```

```{r}
cont.matrix <- makeContrasts(
  time = "time.grp",
  levels = design
)
```

```{r}
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
top_table <- topTable(fit.cont, adjust="fdr", n = Inf)
top_table$Accession <- rownames(top_table)
top_table_2 <- merge(annotationInfo, top_table, by = "Accession")
save_file_excel(top_table_2, "LIMMA_time_series", paste0("results/", name_ana, "/Time_series/"))
top_table_3 <- top_table_2
top_table_3 <-top_table_3[top_table_3$adj.P.Val < 0.05, ]
top_table_3$Significance <- ifelse(top_table_3$logFC < 0, paste0("downregulated"), paste0("upregulated"))
save_file_excel(top_table_3, "LIMMA_time_series_filtered", paste0("results/", name_ana, "/Time_series/"))
```

```{r}
antiviral_database <- read.xlsx("data/AntiViral_DB.xlsx", 3)
```

```{r}
top_table_3 <- top_table_3[top_table_3$Gene_name %in% antiviral_database$Genes,]
```


```{r}
top_table_3 <- top_table_3[top_table_3$Gene_name != "",]
```

```{r}
input$Accession <- rownames(input)
input <- select(input, Accession, contains("UNI"), contains("24"),  contains("48"), contains("72"))
```

```{r}
data_heatmap <- merge(top_table_3, input, by = "Accession")
data_heatmap <- data_heatmap[order(data_heatmap$logFC), ]
```
```{r}
genes <- data_heatmap$Gene_name
```

```{r}
my_palette <- colorRampPalette(c("blue","white","red"))(n=20)
```

```{r}
data_heatmap_2 <- scale(t(data_heatmap[,-c(1:9)]))
```
```{r}
colnames(data_heatmap_2) <- genes
```
```{r}
data_heatmap_2 <- data.frame(t(data_heatmap_2))
```

```{r}
pdf(paste0("results/", name_ana, "/figures/covid-19_proteomics_heatmap.pdf"), height = 20)
heatmap.2(as.matrix(data_heatmap_2), dendrogram = "both", Colv = FALSE, Rowv = FALSE, trace = "none", scale = "none", col = my_palette, key = FALSE, keysize = 0.01, cexCol = 0.7)
dev.off()
```

```{r}
top_table_up <- top_table_3[top_table_3$Significance == "upregulated",]
top_table_down <- top_table_3[top_table_3$Significance == "downregulated",]
```

```{r}
write.table(data.frame(top_table_up$Gene_name), file = paste0("results/", name_ana, "/Time_series/12F_up_regulated.txt"), sep = "\t",
            row.names = FALSE)

write.table(data.frame(data = top_table_down$Gene_name), file = paste0("results/", name_ana, "/Time_series/12F_down_regulated.txt"), sep = "\t",
            row.names = FALSE)
```

```{r}
system("sed 's/\"//g' results/covid-19_proteomics/Time_series/12F_down_regulated.txt > results/covid-19_proteomics/Time_series/12F_down_regulated_2.txt")
```

```{r}
system("sed 's/\"//g' results/covid-19_proteomics/Time_series/12F_up_regulated.txt > results/covid-19_proteomics/Time_series/12F_up_regulated_2.txt")
```

## python code (run in ipython)

import gseapy

gseapy.enrichr(gene_list="results/covid-19_proteomics/Time_series/12F_up_regulated_2.txt",description='12F_up',gene_sets="data/KEGG_advanced/KEGG.gmt",outdir='results/GSEA',cutoff=0.5,verbose=True, background  = 7757)

gseapy.enrichr(gene_list="results/covid-19_proteomics/Time_series/12F_down_regulated_2.txt",description='12F_down',gene_sets="data/KEGG_advanced/KEGG.gmt",outdir='results/GSEA',cutoff=0.5,verbose=True, background  = 7757)


## extract viral proteins
```{r}
list_viral_proteins <- c("P0DTC9", "P0DTC5", "Q9P266", "P0DTC2", "P0DTC3", "P0DTC1")
```

```{r}
data_viral <- input[input$Accession %in% list_viral_proteins,]
rownames(data_viral) <- data_viral$Accession
data_viral$Accession <- NULL
data_viral <- select(data_viral, contains("UNI"), contains("24"),  contains("48"), contains("72"))
```

```{r}
data_viral <- data.frame(t(data_viral))
```

```{r}
input_plot <- data_viral
input_plot$time <- c(rep(0, nb_samples), rep(24, nb_samples), rep(48, nb_samples), rep(72, nb_samples))
input_plot$group <- as.factor(rep(c(1, 2,3), 4))
```
```{r}
library(ggplot2)
```

```{r}
for(i in 1:6){
  pltbase <- ggplot()+ geom_line(data = input_plot, aes(x=time, y=input_plot[,i], group=group), size=2, color = "black", alpha=0.4)+ theme_classic()
  ggsave(paste0("results/", name_ana, "/figures/", "viral_protein_", names(input_plot)[i], ".pdf"))
}
```

