---
title: "Pairwise comparisons proteomics"
output: html_notebook
---


## Data specificity
- paired samples
- time course experiments
- 4 conditions(6 comparisons)


## Data input and output
Inputs : 
Proteomics raw data is submitted with uniprot accession as row and samples as columns.
Annotation file with as columns 1) Uniprot Accession 2) Gene name 

Ouputs :  
1)  PCA 1/ 2, UMAP, PLS-DA, Distribution after norm
2)  Heatmap, pathways, figure with pathways
3)  Scatter plot, Correlation matrix, pathways, figure pathways


## Outline 
### Pre-processing
Count missing values / PCA 1
Normalization ( NormalizerDE, selection based on report)
Plotting Distribution / Dimension reduction : PCA 2 / UMAP / PLS-DA

### Multiple comparisons
LIMMA : 6 comparisons (UNINF/T1 / T2 / T3) paired samples
Heatmap (hierarchical clustering) with proteins significant in at least one infected sample compared to control (pvalue  < 0.05).
Cytoscape Network → Reactome on obtained clusters

### Correlation viral/ host proteins
Scatter plot log2 (viral_protein_inf_sample/viral_prot_control) vs time after infection
Calculate averaged profile of viral proteins
Pearson correlation host proteins/ averaged profile of all viral proteins => select FDR < 0.01
Cytoscape Network → biological process GO terms (plot network)


### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Corona_omics/")) 
```

### load packages
```{r}
library(ggplot2)
library(limma)
library(dplyr)
library(xlsx)
library(RColorBrewer)
library(gplots)
library(ggrepel)
library(knitr)
library(ggfortify)
#library(NormalyzerDE)
opts_chunk$set(echo = TRUE)
options(bitmapType="cairo")
```

### load own functions
```{r}
source("src/usefull_functions.R")
```
```{r}
name_ana <- "covid-19_proteomics"

nb_samples <- 3

# define conditions
t0 <- "UNINF"
t24 <- "H24PI"
t48 <- "H48PI"
t72 <- "H72PI"

# define associated colors
c_t0 <- "#B1A78C"
c_t24 <- "#687E8A"
c_t48 <- "#41607D"
c_t72 <- "#09253B"
  
condition <- as.factor(c(rep(t0, 2), rep(t24, nb_samples), rep(t48, nb_samples), rep(t72, nb_samples)))
colors <- as.factor(c(rep(c_t0, nb_samples), rep(c_t24, nb_samples), rep(c_t48, nb_samples), rep(c_t72, nb_samples)))
gg_colors <- c(c_t24,c_t48, c_t72, c_t0)

input <- read.delim("processing/norm_files/covid-19_proteomics_norm_file.txt")
rownames(input) <- input$X
input$X <- NULL

annotationInfo <- read.csv("processing/covid-19_proteomics/covid-19_proteomics_annotationInfo.csv")
annotationInfo$Gene_name <- gsub(";.*","",annotationInfo$Gene_name)
annotationInfo$X <- NULL
```

## remove outlier due to technical error
```{r}
input$UNINF_3 <- NULL
```

## Multiple comparisons

## four time points experiments = 

## set levels
```{r}
groups <- as.factor(condition)
design <- model.matrix( ~ 0 + groups)
colnames(design) <- levels(groups)
```

## fit model
```{r}
fit <- lmFit(input, design)
```


```{r}
list_comp <- c("H24PI-UNINF", "H48PI-UNINF","H72PI-UNINF","H48PI-H24PI", "H72PI-H24PI","H72PI-H48PI")
```

```{r}
DGE_results <- data.frame(annotationInfo)
DGE_results$X <- NULL
```

```{r}
antiviral_database <- read.xlsx("data/AntiViral_DB.xlsx", 3)
```
```{r}
for(comp in list_comp){
  c2 <- substr(comp, 1,5)[1]
  c1 <- substr(comp, 7,11)[1]
  dge <- extract_limma(fit, design, name_ana, c1, c2)
  print(dge)
  dge <- dge[dge$Gene_name %in% antiviral_database$Genes, ]
  top_table_up <- dge[dge$Significance_2 == "upregulated",]
  top_table_down <- dge[dge$Significance_2 == "downregulated",]
  print(top_table_up)
  name_up <- paste0("results/",name_ana,"/Pathways_heatmap/prot_up_regulated_", comp, ".txt")
  name_down <- paste0("results/",name_ana,"/Pathways_heatmap/prot_down_regulated_", comp, ".txt")
  print(name_up)
  print(name_down)
  write.table(data.frame(top_table_up$Gene_name), file = name_up, sep = "\t",
            row.names = FALSE)
  write.table(data.frame(data = top_table_down$Gene_name), file = name_down, sep = "\t",
            row.names = FALSE)
}
```


```{r}
system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H24PI-UNINF.txt > results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H24PI-UNINF_2.txt")

system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H24PI-UNINF.txt > results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H24PI-UNINF_2.txt")
```

```{r}
system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H48PI-H24PI.txt > results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H48PI-H24PI_2.txt")

system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H48PI-H24PI.txt > results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H48PI-H24PI_2.txt")
```

```{r}
system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H72PI-H48PI.txt > results/covid-19_proteomics/Pathways_heatmap/prot_up_regulated_H72PI-H48PI_2.txt")

system("sed 's/\"//g' results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H72PI-H48PI.txt > results/covid-19_proteomics/Pathways_heatmap/prot_down_regulated_H72PI-H48PI_2.txt")
```


